Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2013-12-29 acivs/acivs.tex
cleanup

%  article.tex (Version 3.3, released 19 January 2008)
%  Article to demonstrate format for SPIE Proceedings
%  Special instructions are included in this file after the
%  symbol %>>>>
%  Numerous commands are commented out, but included to show how
%  to effect various options, e.g., to print page numbers, etc.
%  This LaTeX source file is composed for LaTeX2e.

%  The following commands have been added in the SPIE class 
%  file (spie.cls) and will not be understood in other classes:
%  \supit{}, \authorinfo{}, \skiplinehalf, \keywords{}
%  The bibliography style file is called spiebib.bst, 
%  which replaces the standard style unstr.bst.

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{float}
\usepackage{multirow}
\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

@@

\author{Peter Cheng, Michael Anderson, Stewart He, and Avideh Zakhor
\skiplinehalf
University of California, Berkeley\\
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photo-realistic visualizations of the data
  collected by such modeling systems. While data acquisition times for
  mobile mapping systems are considerably shorter than for static
  ones, their recovered camera poses often suffer from inaccuracies,
  resulting in visible discontinuities when successive images are
  projected onto a surface for texturing. We present a method for
  texture mapping that starts by selecting images whose camera poses
  are well-aligned in two dimensions. We then align images to geometry
  as well as to each other, producing visually consistent textures
  even in the presence of inaccurate surface geometry and noisy camera
  poses. The final step is to select and composite images into a
  texture mosaic. The effectiveness of the proposed method is
  demonstrated on a number of different indoor environments.
\end{abstract}

@@

\keywords{Texture Mapping, Reconstruction, Image Stitching, Mosaicing}

@@

In this paper, we present an approach to texture mapping 3D models of
indoor environments made of planar surfaces in the presence of
uncertainty and noise in camera poses. In particular, we consider data
obtained from a human-operated backpack system with a number of laser
range scanners as well as 2 cameras facing left and right, each
equipped with fisheye lenses reaching an approximately 180$^{\circ}$
field of view and taking photos at a rate of 5 Hz. Applying multiple
localization and loop-closure algorithms on the raw data collected by
the onboard sensors \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}, the backpack is localized\footnote{In this paper, we
  use the terms localization and pose recovery interchangeably, in
  that they both refer to recovering position and orientation.}  over
its data collection period. This requires recovering the 6 degrees of
freedom for the backpack itself as well as the cameras rigidly mounted
on it. Once this is complete, the data from the laser range scanners
is used to generate a 3D point cloud of the surrounding environment,
from which a 3D planar model is created \cite{sanchez2012point}. This
model, consisting of 2D polygonal planes in 3D space, along with the
set of images captured by the backpack's cameras and their noisy 3D
poses, can be considered the input to our texture mapping problem.

\begin{figure}
  \centering
  \includegraphics[width=6in]{flowchart.jpg}
  \caption{The proposed texture mapping procedure\\}
  \label{fig:flowchart}
\end{figure}

The overall block diagram for the proposed texture mapping procedure
is shown in Figure \ref{fig:flowchart}. We texture map each planar
surface independently and in parallel. For each surface, we begin by
obtaining a set of images that spans the entire surface with high
resolution imagery. We then use our noisy camera poses to project
these selected images onto the surface. These projections are then
refined in 2D, in order to maximally align them with the surface's
geometry, as well as to each other, allowing us to handle both errors
in geometry as well as camera poses. For surfaces and camera poses at
arbitrary locations or orientations, generally ceilings and floors, we
propose a tile-based approach for sampling high-resolution portions of
images and compositing them into a texture. In cases where cameras
have consistently perpendicular viewing angles to the surfaces under
consideration, generally walls, we demonstrate a superior approach
that leads to more seamless textures.

@@

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
There are many existing approaches to stitching together multiple
images to produce a larger, seamless image \cite{szeliski2006image,
  agarwalapanoramas, wangmultipleviews, coorg1997matching,
  debevechybrid, bernardinimultiplescans}. Generally, parts of images
are matched to each other, either through direct comparisons or
feature detection and matching. Images are then transformed to
maximize matches, often by computing homographies between pairs of
images, or by iteratively adjusting camera poses in 1 to 6 degrees of
freedom.

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar environments, and
rotational differences between camera poses
\cite{szeliski2006image}. Feature matching however, works best when
multiple unique visual references exist in the environment that can be
detected in multiple images. In contrast, indoor environments have a
high prevalence of bare surfaces, as well as repeating textures, such
as similar windows, doors, and wall-mounted decorations, that are
difficult to disambiguate. This lack of reliable reference points
often results in errors when matching images together.

Additionally, our datasets often contain long chains of images,
corresponding to long hallways and corridors as shown in Figure
\ref{fig:mosaic3D}, which leads to error accumulation when image
correspondences are not accurate. For example, when matching a long
chain of images through homography, a pixel in the $nth$ image must be
translated into the first image's coordinates by multiplying by the
$3\times3$ matrix $H_1 H_2 H_3 ... H_n$. Any error in one of these
homography matrices is propagated to all further images, resulting in
drift.

\begin{figure}
  \centering \subfloat[][]{\includegraphics[trim=0cm 0cm 20cm 0cm,
    clip=true, width=6in, height=0.7in]{graphApproach.jpg}}

  \centering \subfloat[][]{\includegraphics[trim=0cm 0cm 15cm 0cm,
    clip=true, width=6in, height=0.7in]{autostitchResult.jpg}}

  \centering \subfloat[][]{\includegraphics[trim=0cm 0cm 75cm 0cm,
    clip=true, width=6in, height=0.7in]{finalLong.jpg}}

  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm, (b) the AutoStitch software package, and (c)
    the proposed method.}
  \label{fig:mosaic3D}
\end{figure}

In prior work, we have experimented with integrating image stitching
with the iterative global localization algorithm used to recover the 6
degrees of freedom of the backpack acquisition system
\cite{liu2010indoor}. When run on long chains of images, especially
where features are sparse, this approach produces distorted textures,
as seen in Figure \ref{fig:mosaic3D}(a). Furthermore, this approach is
not closed-form, and its iterative camera adjustment process over
large datasets led to prohibitively long computation time.

The AutoStitch software package performs homography-based alignment as
well, with additional provisions that attempt to reduce drift and
increase efficiency \cite{panorama2d, autostitch}. Though AutoStitch
performs well in areas with dense features, it can not handle areas
without features, and has trouble aligning wall sections with even
short segments of bare texture. The example in Figure
\ref{fig:mosaic3D}(b) has been generated after many rounds of manual
tuning; we have empirically found that areas with fewer visual
features or repeating texture patterns simply failed outright with
AutoStitch.

@@

The geometry of the texture mapping process for a planar surface is
shown in Figure \ref{fig:projection}. Given a set of $M$ images to
texture a target plane, camera matrix $P_i$ for the $ith$ image
transforms a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. A camera matrix $P_i$ is composed of
the camera's intrinsic parameters, containing focal length and image
center, as well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and the corresponding localization
algorithms \cite{chen2010indoor, liu2010indoor, kua2012loopclosure}
and are noisy.

Since the backpack system takes pictures at a rate of 5 Hz, hundreds
of images are available for texturing each surface in the 3D
model. Our objective in designing a texture mapping process is to
determine which of these images should be used, and where their
contents should map onto the final texture, in order to minimize any
visual discontinuities or seams that would suggest that the plane's
texture is not composed of a single continuous image. In the remainder
of this section, we propose an image selection criteria and use it to
derive a simple tile-based texture mapping procedure.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.8in]{Projection.pdf}
    \caption{Surfaces to be textured are specified in 3D space by
      corners $C_i$. Images are related to each surface through the
      camera matrices $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.8in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
a straightforward approach for texturing is to discretize the target
plane into small square tiles, and choose an image to texture each
tile directly. We choose to work with rectangular units to ensure that
borders between any two distinct images in the final texture are
either horizontal or vertical. Since most environmental features
inside buildings are horizontal or vertical, any visible seams in the
texture intersect them minimally and are therefore less noticeable.

In order to select an image for texturing a tile $t$, we first gather
a list of candidate images that contain all four of its corners; we
can rapidly check this by projecting $t$ into each image using the
$P_i$ camera matrices. Furthermore, each candidate image must have
been taken at a time when its camera had a clear line-of-sight to $t$,
which can be determined using standard ray-polygon intersection tests
between the camera location, $t$, and every other surface
\cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since resolution decreases and camera pose errors become more
pronounced with distance, we wish to minimize the distance between
cameras and the surfaces they texture. Additionally, we desire images
that are projected perpendicularly, rather than obliquely, onto the
plane, maximizing the resolution and amount of useful texture
available in their projections, as well as minimizing any parallax
effects due to real-world geometry not accurately represented by the
digital 3D model. In other words, we wish to minimize the angle
between the tile's normal vector and the camera axis for images
selected for texturing that tile. These two criteria can be met by
maximizing the function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$
as shown in Figure \ref{fig:scoringFunction}, where $d$ is the
distance between the centers of a camera and a tile, and $\vec{n}$ and
$\vec{c}$ are the directions of the plane's normal and the camera axis
respectively. As Figure \ref{fig:compareAll}(a) demonstrates, this
approach, when used directly for texturing, results in image
boundaries with abrupt discontinuities between tiles, due to
misalignment between images, though most of it appears to be
reconcilable using 2D transforms.

While this procedure can be used to generate a texture, it is also
useful for obtaining a collection of images that can be used as a
starting point for our proposed texturing procedure, as explained in
Section \ref{sec:2dAlignment}. For our datasets, this process
generally selects around 10\% of all the images that could be used for
texturing the surface. This roughly 10:1 subsampling not only reduces
computational complexity of the remaining steps, but also selects the
most promising images for the remaining steps.

Given the large number of input images, it is possible to develop a
more sophisticated image selection procedure, as described in Section
\ref{sec:imageCompositing}, however, a more significant and reliable
improvement can be found through first improving the alignment of the
images as shown in the next section.

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift_blend.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=3.4in]{wall2_cache_comparison.png}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=3.4in]{wall2_shortest_comparison.png}}

@@

\section{2D Image Alignment}
\label{sec:2dAlignment}
In this section, we describe our method for efficient and robust image
alignment. Rather than register all of our images in 3D, as many
state-of-the-art techniques for image stitching do, we instead select
a subset of images to be aligned in 2D. This subset corresponds to the
images selected by the image selection procedure described in Section
\ref{sec:simpleTextureMapping}.

Applying 2D alignments to this set of images works well for a number
of reasons. First, the nature of our input data and the selected
images is such that localization error chiefly occurs in two
dimensions, which correspond to the plane of the surface being
projected onto. This is because the backpack operator, during data
acquisition, makes efforts to walk within a few meters of, and
parallel to all walls being scanned. As a result, the translational
error of camera poses is quite minimal in the dimension perpendicular
to the wall surface being textured. Furthermore, because our set of
images has minimal distance from its corresponding wall surface,
rotational errors are less pronounced, except for rotation around the
axis perpendicular to the surface being textured. Therefore, most
translation and rotational errors occur in the 2 dimensions of the
wall surface's plane. This of course does not apply for ceilings and
floors; however, the lack of features on most ceilings and floors
means that 3D misalignment is not likely to visually make a
difference, beyond what can be done in 2D.

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=6in]{allunshifted.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{allshifted.jpg}}

  \centering \subfloat[][]{\includegraphics[trim=0cm 2cm 0cm 0cm,
    clip=true, width=5in]{matchLines.jpg}}

  \caption{Images projected onto a ceiling surface, where
    geometry-based lines corresponding to the ceiling's boundary are
    shown in red. Image-based lines detected by Hough transform in the
    image projections are shown in blue; (a) images projected with
    their original noisy camera poses; (b) after image alignment to
    maximize line matches between images and geometry; (c) examples of
    matching lines in cases with $\geq$ 2 line pairs, 1 line pair, and
    zero line pairs, from top to bottom.}
  \label{fig:geometryAlignment}
\end{figure}

After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we obtain a set
of image-based line segments by using Hough transforms to detect lines
in the image projections. We then gather a set of geometry-based
lines, which correspond to the lines comprising the target surface's
border, as well as lines formed where other surfaces intersect the
target surface. An example of these lines is shown in red for a
ceiling surface in Figure \ref{fig:geometryAlignment}(a). Ideally, for
perfect camera poses and surface geometry, the lines in images
corresponding to corners between surfaces should match up exactly with
corners in the 3D model. By inducing such lines to match, we fit
camera poses more accurately to the surface, and to each other as
well.

To align images to surface geometry, we collect pairs of image-based
and geometry-based line segments, which are within a distance and
angular difference threshold of each other. We have found a distance
threshold of 250 mm and an orientation difference threshold of
$10^\circ$ to work well for our datasets. For each pair of lines, we
compute the angular difference between the pair's image and geometry
lines. If there are 2 or more pairs with angular differences within
$1^\circ$, we select the two pairs with the longest noncollinear
image-based lines, and rotate the image such that the lines in the
pair with the longest image-based line become parallel. We then find a
translation such that the lines in that same pair overlap. This
translation has ambiguity in the dimension along the matched lines,
which is resolved by matching the midpoint of the image-based line in
the second pair to its corresponding geometry-based line. This
corresponds to case (1) of Figure
\ref{fig:geometryAlignment}(c). Thus, with 2 or more pairs, it is
possible to obtain a fixed rotation and translation, which are saved
for usage in Section \ref{sec:robustSIFTFeatureMatching}.

If there are not 2 or more pairs with similar angular differences, we
select the pair with the longest image-based line and apply a rotation
and the minimal translation to match its lines. This translation's
ambiguity however, can not be resolved, but is also saved to be used
in Section \ref{sec:robustSIFTFeatureMatching}. This corresponds to
case (2) of Figure \ref{fig:geometryAlignment}(c). Finally, in the
case where there are no line pairs, we can still rotate images in
order to exploit patterns in indoor environments (case (3) of Figure
\ref{fig:geometryAlignment}(c)). For instance, doors, windows,
furniture, etc. tend to have linear edges that are parallel to the
edges of the surfaces they are on. Similarly, lights, visible interior
walls, etc. which are visible in floor and ceiling images, tend to be
parallel to exterior walls corresponding to the edges of ceiling and
floor surfaces. Thus, we choose to minimize the angle between
image-based lines and geometry-based lines. We use the RANSAC
framework, with an inlier difference of $0.5^\circ$ in order to
successfully ignore outliers \cite{fischler1981random}.

After these steps, image projections line up well with target
surfaces, as shown in Figure \ref{fig:geometryAlignment}(b), which is
considerably more aligned than Figure \ref{fig:geometryAlignment}(a)
is. This procedure reconciles both errors in camera poses as well as
in geometry, and results in sharp, continuous borders across images,
which is crucial when checking for occlusion.

@@

\begin{figure}
  \centering
  \begin{tabular}{cc}
    \subfloat[][]{\raisebox{-1in}{\includegraphics[width=3in]{occlusiondiagram.png}}} &

  \begin{tabular}{c}
    \subfloat[][]{\includegraphics[trim=0cm 0cm 1.3cm 0cm,
    clip=true, width=3in,
      height=0.9in]{occlusionimagebad.png}} \\
    \subfloat[][]{\includegraphics[trim=1.3cm 0cm 0cm 0cm,
    clip=true, width=3in,
      height=0.9in]{occlusionimagegood.png}}
  \end{tabular}
\end{tabular}
\caption{(a) The image from the camera in this diagram contains
  texture that belongs to the gray occluding surface, which should not
  be projected onto the orange target surface; (b) without geometry
  alignment, texture to the left of the red line would be removed,
  which would leave some erroneous texture projected onto our target
  surface; (c) after geometry alignment, the image is shifted,
  resulting in the correct amount of texture being removed.}
\label{fig:occlusion}
\end{figure}

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a gray
occluding surface. We remove this incorrect texture by recursively
performing ray-polygon intersection tests between the camera location
and other surfaces \cite{rayintersection}. These intersection tests
are performed at the corners of a grid overlaid upon the target
surface. Where all four corners of a grid section are occluded,
texture is removed. Where one or more corners are occluded, the grid
is subdivided into four, and the process repeats. Occlusion checking
works entirely with geometry, so by ensuring that images match
geometry using \ref{sec:geometryAlignment}'s alignment procedure,
texture belonging to other surfaces is accurately removed, as seen in
Figure \ref{fig:occlusion}(b).

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
The next step is to align the selected images from Section
\ref{sec:simpleTextureMapping} for each surface to each other by
searching for corresponding feature points between all pairs of
overlapping images. We use feature alignment rather than pixel or
intensity-based alignment due to the differences in lighting as well
as possible occlusion among images, both of which feature alignment is
less sensitive to \cite{lowe1999object, mikolajczyk2005performance,
  szeliski2006image}. We use SiftGPU \cite{siftgpu} for its high
performance on both feature detection as well as pairwise
matching. These matches determine $dx$ and $dy$ distances between each
pair of features for two image projections, though these distances may
not always be the same for different features. Since indoor
environments often contain repetitive features such as floor tiles or
doors, we need to ensure that SIFT-based distances are
reliable. First, we only align parts of images that overlap given the
original noisy poses. Second, we discard feature matches that
correspond to an image distance greater than 200 mm from what the
noisy poses estimate. In order to utilize the remaining feature
matches robustly, RANSAC \cite{fischler1981random} is again used to
estimate the optimal $dx_{i,j}$ and $dy_{i,j}$ distances between two
images $i$ and $j$. We use a 5 mm threshold for RANSAC, so that SIFT
matches are labeled as outliers if their distance is not within 5 mm
of the sampled average distance.

>>>>

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{float}
\usepackage{multirow}
\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

@@

\author{Peter Cheng \and Michael Anderson \and Stewart He \and Avideh Zakhor}
\institute{University of California, Berkeley\\ Berkeley CA 94720, USA}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photo-realistic visualizations of the data
  collected by such modeling systems. While data acquisition times for
  mobile mapping systems are considerably shorter than for static
  ones, their recovered camera poses often suffer from inaccuracies,
  resulting in visible discontinuities when successive images are
  projected onto a surface for texturing. We present a method for
  texture mapping that starts by selecting images whose camera poses
  are well-aligned in two dimensions. We then align images to geometry
  as well as to each other, producing visually consistent textures
  even in the presence of inaccurate surface geometry and noisy camera
  poses. Images are then composited into a final texture mosaic. The
  effectiveness of the proposed method is demonstrated on a number of
  different indoor environments.

  \keywords{Texture Mapping, Reconstruction, Image Stitching, Mosaicing}
\end{abstract}

@@

@@

In this paper, we present an approach to texture mapping 3D models of
indoor environments made of planar surfaces in the presence of
uncertainty and noise in camera poses. In particular, we consider data
obtained from a human-operated backpack system with a number of laser
range scanners as well as 2 cameras facing left and right, each
equipped with fisheye lenses reaching an approximately 180$^{\circ}$
field of view and taking photos at a rate of 5 Hz
\cite{liu2010indoor}. Applying multiple localization and loop-closure
algorithms on the raw data collected by the onboard sensors
\cite{chen2010indoor, kua2012loopclosure, liu2010indoor}, the backpack
is localized\footnote{In this paper, we use the terms localization and
  pose recovery interchangeably, in that they both refer to recovering
  position and orientation.}  over its data collection period. This
involves recovering the 6 degrees of freedom for the backpack itself
as well as the cameras rigidly mounted on it. Once this is complete,
the data from the laser range scanners is used to generate a 3D point
cloud of the surrounding environment, from which a 3D planar model is
created \cite{sanchez2012point}. This model, consisting of 2D
polygonal planes in 3D space, along with the set of images captured by
the backpack's cameras and their noisy 3D poses, can be considered the
input to our texture mapping problem.

\begin{figure}
  \centering
  \includegraphics[width=4.5in]{flowchart.jpg}
  \caption{The proposed texture mapping procedure\\}
  \label{fig:flowchart}
\end{figure}

The overall block diagram for the proposed texture mapping procedure
is shown in Figure \ref{fig:flowchart}, where the number attached
attached to each box indicates the section in which the concept in the
box is explained in this paper. We texture map each planar surface
independently and in parallel. For each surface, we begin by selecting
a set of images that spans the entire surface with high resolution
imagery. We then use our noisy camera poses to project these selected
images onto the surface. These projections are then refined in 2D, in
order to maximally align them with the surface's geometry, as well as
to each other, allowing us to handle both errors in geometry as well
as camera poses. For surfaces and camera poses at arbitrary locations
or orientations, generally ceilings and floors, we propose a
tile-based approach for sampling high-resolution portions of images
and compositing them into a texture. In cases where cameras have
consistently perpendicular viewing angles to the surfaces under
consideration, generally walls, we demonstrate a superior approach
that leads to more seamless textures.

@@

@@

% The geometry of the texture mapping process for a planar surface is
% shown in Figure \ref{fig:projection}. Given a set of $M$ images to
% texture a target plane, camera matrix $P_i$ for the $ith$ image
% transforms a 3D point in the world coordinate system to a 2D point or
% pixel in image $i$'s coordinates. A camera matrix $P_i$ is composed of
% the camera's intrinsic parameters, containing focal length and image
% center, as well as extrinsic parameters which specify the rotation and
% translation of the camera's position in 3D world coordinates at the
% time that image $i$ is taken. These extrinsic parameters are
% determined by the backpack hardware and the corresponding localization
% algorithms \cite{chen2010indoor, liu2010indoor, kua2012loopclosure}
% and are noisy.

Since the backpack system takes pictures at a rate of 5 Hz, hundreds
of images are available for texturing each surface in the 3D model. To
obtain a selection of images to work with, we discretize each surface
into small square tiles, and for each tile, gather the images with a
clear line of sight to it. We then find the image that maximizes the
scoring function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$ for
each tile, as shown in Figure \ref{fig:scoringFunction}, where $d$ is
the distance between the centers of a camera and a tile, and $\vec{n}$
and $\vec{c}$ are the directions of the plane's normal and the camera
axis respectively. In Figure \ref{fig:compareAll}(a), an image has
been textured using the images selected for each tile. While images
with good viewing angles and high resolution have been used, there are
a significant amount of discontinuities, though many appear
reconcilable with 2D transformations.

By aggregating the images selected for tiles on each surface, we
obtain a collection of images that can be used as a starting point for
the proposed texturing procedure, as explained in Section
\ref{sec:2dAlignment}. For our datasets, this process generally
selects around 10\% of all the images that could be used for texturing
the surface. This roughly 10:1 subsampling removes images with poor
viewing angles, ensuring that the most promising images are considered
in the following steps.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.2in,width=2in]{Projection.pdf}
    \caption{Surfaces to be textured are specified in 3D space by
      corners $C_i$. Images are related to each surface through the
      camera matrices $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.2in, width=2in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_naive_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_cache_shift.jpg}}
  \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_shortest_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_shortest_shift_blend.pdf}}

  \centering \subfloat[][]{\includegraphics[width=2.4in,
    height=2.2in]{wall2_cache_comparison.png}}
  \subfloat[][]{\includegraphics[width=2.4in,
    height=2.2in]{wall2_shortest_comparison.png}}

@@

\section{2D Image Alignment}
\label{sec:2dAlignment}
% In this section, we describe our method for efficient and robust image
% alignment. Rather than register all of our images in 3D, as many
% state-of-the-art techniques for image stitching do, we instead select
% a subset of images to be aligned in 2D. This subset corresponds to the
% images selected by the image selection procedure described in Section
% \ref{sec:simpleTextureMapping}.

% Applying 2D alignments to this set of images works well for a number
% of reasons. First, the nature of our input data and the selected
% images is such that localization error chiefly occurs in two
% dimensions, which correspond to the plane of the surface being
% projected onto. This is because the backpack operator, during data
% acquisition, makes efforts to walk within a few meters of, and
% parallel to all walls being scanned. As a result, the translational
% error of camera poses is quite minimal in the dimension perpendicular
% to the wall surface being textured. Furthermore, because our set of
% images has minimal distance from its corresponding wall surface,
% rotational errors are less pronounced, except for rotation around the
% axis perpendicular to the surface being textured. Therefore, most
% translation and rotational errors occur in the 2 dimensions of the
% wall surface's plane. This of course does not apply for ceilings and
% floors; however, the lack of features on most ceilings and floors
% means that 3D misalignment is not likely to visually make a
% difference, beyond what can be done in 2D.

@@

\begin{figure}
  \centering
  \begin{tabular}{cc}
    \subfloat[][]{\raisebox{-0.8in}{\includegraphics[trim=0cm 2cm 0cm 0cm,
        clip=true, width=2.4in]{matchLines.jpg}}} &
    \begin{tabular}{c}
      \subfloat[][]{\includegraphics[width=2.4in]{allunshifted.jpg}} \\
      \subfloat[][]{\includegraphics[width=2.4in]{allshifted.jpg}}
    \end{tabular}
  \end{tabular}
  \caption{Images projected onto a ceiling surface, where
    geometry-based lines corresponding to the ceiling's boundary are
    shown in red. Image-based lines detected by Hough transform in the
    image projections are shown in blue; (a) examples of matching
    lines in cases with $\geq$ 2 line pairs, 1 line pair, and zero
    line pairs, from top to bottom; (b) images projected with their
    original noisy camera poses; (c) after image alignment to maximize
    line matches between images and geometry; }
  \label{fig:geometryAlignment}
\end{figure}

To align images to geometry, line segments are detected in images via
Hough transforms. We also gather a set of geometry-based lines, which
correspond to lines comprising the target surface's border, as well as
lines formed where other surfaces intersect the target surface. An
example of these lines is shown in red for a ceiling surface in Figure
\ref{fig:geometryAlignment}(a). Ideally, for perfect camera poses and
surface geometry, the lines in images corresponding to corners between
surfaces should match up exactly with corners in the 3D model. By
inducing such lines to match, we fit camera poses more accurately to
the surface, and therefore to each other as well.

First, pairs of image-based and geometry-based line segments within a
distance and angular difference threshold of each other are
collected. We have found a distance threshold of 250 mm and an
orientation difference threshold of $10^\circ$ to work well for our
datasets. Each such pair signifies a rotation and translation with one
degree of freedom that matches an image line to a geometry line. Pairs
are then binned by rotation in $1^\circ$ increments, which reveals the
rotation that maximizes line matches. From the pairs with that
corresponding rotation, the two longest noncollinear image-based lines
are used to compute a fixed translation. An example of this is case
(1) of Figure \ref{fig:geometryAlignment}(c). If a rotation is
calculated, but a fixed translation cannot be obtained, the pair with
the longest image-based line is selected, and the rotation and the
minimal translation to match its lines are applied. This corresponds
to case (2) of Figure \ref{fig:geometryAlignment}(c). Finally, if no
line pairs were detected, images are rotated using a RANSAC
\cite{fischler1981random} approach to maximize parallel image and
geometry lines, without regard for translational distance. This
corresponds to case (3) of Figure \ref{geometryAlignment}(c). The
results of this procedure are recorded, as the translational degrees
of freedom for each image are used in Section
\ref{sec:robustSIFTFeatureMatching}.

After this step, image projections line up well with target surfaces,
as shown in Figure \ref{fig:geometryAlignment}(b), which is
considerably more aligned than Figure
\ref{fig:geometryAlignment}(a). This procedure reconciles both errors
in camera poses as well as in geometry, and results in sharp,
continuous borders across images, which is crucial when checking for
occlusion.

@@

\begin{figure}
  \centering
  \begin{tabular}{cc}
    \subfloat[][]{\raisebox{-0.8in}{\includegraphics[width=2.4in]{occlusiondiagram.png}}} &

    \begin{tabular}{c}
      \subfloat[][]{\includegraphics[trim=0cm 0cm 1.3cm 0cm,
        clip=true, width=2.4in,
        height=0.6in]{occlusionimagebad.png}} \\
      \subfloat[][]{\includegraphics[trim=1.3cm 0cm 0cm 0cm,
        clip=true, width=2.4in,
        height=0.6in]{occlusionimagegood.png}}
    \end{tabular}
  \end{tabular}
\caption{(a) The image from the camera in this diagram contains
  texture that belongs to the gray occluding surface, which should not
  be projected onto the orange target surface; (b) without geometry
  alignment, texture to the left of the red line would be removed,
  which would leave some erroneous texture projected onto our target
  surface; (c) after geometry alignment, the image is shifted,
  resulting in the correct amount of texture being removed.}
\label{fig:occlusion}
\end{figure}

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a gray
occluding surface. We remove this incorrect texture by recursively
performing ray-polygon intersection tests between the ray from the
camera location to locations on the target surface
\cite{rayintersection}. These intersection tests are performed at the
corners of a grid overlaid upon the target surface. Where all four
corners of a grid section are occluded, texture is removed. Where one
or more corners are occluded, the grid is subdivided into four, and
the process repeats. Occlusion checking works entirely with geometry,
so by ensuring that images match geometry using
\ref{sec:geometryAlignment}'s alignment procedure, texture belonging
to other surfaces is accurately removed, as seen in Figure
\ref{fig:occlusion}(b).

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
The next step is to align images to each other by detecting feature
point matches between all pairs of overlapping images. We use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among images,
both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance,
  szeliski2006image}. We use SiftGPU \cite{siftgpu} for its high
performance on both feature detection as well as pairwise
matching. Detected matches determine $dx$ and $dy$ distances between
each pair of features for two image projections, though distances
often are not the same for different features. Since indoor
environments often contain repetitive features such as floor tiles or
doors, we need to ensure that SIFT-based distances are
reliable. First, we only detect feature matches in the parts of images
that overlap given the original noisy poses. Second, feature matches
that correspond to an image distance greater than 200 mm from what the
noisy poses estimate are discarded. In order to utilize the remaining
feature matches robustly, RANSAC \cite{fischler1981random} is used to
estimate the optimal $dx_{i,j}$ and $dy_{i,j}$ distances between two
images $i$ and $j$.
